---
title: "Lab1_Logistic"
author: "Majbritt Lindholt Stenbøg"

Used packages: 
```{r}
library(apastats) # For reporting in APA
library(dplyr)
library(tidyverse)
library(pscl) 
library(lmtest) 	
library(questionr) 
library(ggplot2)
```

At first I load my data and view the table, and remove the NAs from my dataset. This is because i am using age and this have 177 NAs
```{r}
titanic <- read.table("Titanic - training set.csv",header=TRUE,sep=",")

view(tidtanic)

titanic_na <- drop_na(titanic)
```

Then I would like to recode my dataset for easier interpretation.
```{r}
# Recoding
titanic_na <- titanic_na %>%
  mutate(Sex = factor(recode(Sex, "0" = "male", "1" = "female")),
         Pclass=factor(recode(Pclass, "1" = "1st", "2" = "2nd", "3"= "3rd")),
         Survived=factor(recode(Survived, "0"="no", "1"="yes")),
         Embarked=factor(recode(Embarked, "C"="Cherbourg", "Q"="Queenstown", "S"="Southampton")))

# Creating a new variable i call Adult_male for males 18 or older
titanic_na <- titanic_na %>% 
  mutate(Adult_male= case_when( (Age >= 18) & 
  (Sex == "male") ~ 1, 
  TRUE ~ 0))
```

In the following, i am including different visualizations and descriptives of my variables.
```{r}
titanic_na %>%
  summary()
  
describe(titanic_na)

#visualise age and sex
fig_1 = titanic_na %>%
  ggplot() +
  aes(y = Age, x = Sex) +
  geom_boxplot()+
  ylim(c(0, 100))

grid.arrange(fig_1)

# Survived
titanic_na %>%
  ggplot() +
  aes(x = Survived) +
  geom_bar()

# Pclass
titanic %>%
  ggplot() +
  aes(x = Pclass) +
  geom_bar()

# Age
titanic %>%
  ggplot() +
  aes(x = Age) +
  geom_histogram()

# Sibsp
titanic %>%
  ggplot() +
  aes(x = SibSp) +
  geom_histogram()

# Survived by sex
barplot(with(titanic_na, table(Survived,Sex)),
        main="Plot of Survived by Sex",
        beside=T,
        ylab="Survived", legend= c("Died", "Survived"),
        xlab="Sex", names = c("Female","Male"))

table(titanic_na$Survived, titanic_na$Sex)
```

In the following, I check for the values of Parent, childen, and siblings. 
```{r}
unique(titanic_na$Parch)
unique(titanic_na$SibSp)
```

Then i tried to create a model 1
```{r}
# prediction based on Parch:
model1 <-  glm(Survived ~ Parch
               , family=binomial, data = titanic)

summary(model1)

predict(model1)
```

And a model 2 - this is the final model
```{r}
finalmodel <-  glm(Survived~ Sex + Age + Pclass + Parch + Adult_male + SibSp,
    family=binomial, 
    data=titanic_na)

summary(finalmodel)

# For APA formatting
describe.glm(finalmodel)
```

A function for turning logits 2 probability, which was from the mentioned source:
```{r}
convert2prob <- function(logit){
  odds <- exp(logit)
  prob <- odds / (1 + odds)
  return(prob)
}
source("https://sebastiansauer.github.io/Rcode/logit2prob.R")
```

Converting to probability:
```{r}
# obtaining the coefficients of the final model
coef(finalmodel)

# Convert the coefficients to probability
convert2prob(coef(finalmodel))

logit_survive <-  5.61968097 + 3* -1.31600276
convert2prob(logit_survive)

predict(finalmodel,
    data.frame(Sex="female", Age=20,Pclass="3rd", Parch=2,Adult_male=1,SibSp=2),
    type = "response")
```

As the model is presented in log(odds) it is hard to intperpret, i therefore use the predict function:
```{r}
predict(finalmodel)
```

A null model for testing the model:
```{r}
null_model <- glm(Survived~1, 
    family=binomial, 
    data= titanic_na)

summary(null_model)
```

create a model without the father in (aduls_male not used compared to the final model)
```{r}
model_nofather <-  glm(Survived ~ Sex + Age + Pclass + Parch+SibSp, 
    family = binomial, 
    data=titanic_na)

summary(model_nofather)
```

likelihood ratio test and AIC - comparing the null model to the final model
```{r}
# Likelihood ratio test
lrtest(null_model,finalmodel)

#AIC test
AIC(null_model,finalmodel)

#Confidence intervals
confint(finalmodel, level=0.95)

# Odds ratio of the final model
odds.ratio(finalmodel)
```

Kate and Sue chance of survival
```{r}
# Kate alone
Kate <- data.frame(Sex="female",
  Age=20,
  Pclass="3rd",
  Parch=1, 
  Adult_male=0,
  SibSp=0)

# Kate with Leonardo
Kate_with_Leo <- data.frame(Sex="female",
  Age=20,
  Pclass="3rd",
  Parch=1, 
  Adult_male=0,
  SibSp=1)

# Kate odds with and without Leonardo
Kate_odds <- exp(predict(finalmodel,Kate))
Kate_odds_Leo <-  exp(predict(finalmodel,Kate_with_Leo))

# Kate probability with and without Leo
Kate_prob <- Kate_odds / (1+Kate_odds)
Kate_prob_Leo <- Kate_odds_Leo / (1+ Kate_odds_Leo)

# Result
summary(Kate_prob)
summary(Kate_prob_Leo)

# Sue alone
Sue <- data.frame(Sex="female",
  Age=4,
  Pclass="3rd",
  Parch=1, 
  Adult_male=0,
  SibSp=0)
  
# Sue with Leonardo
Sue_with_Leo <- data.frame(Sex="female",
  Age=4,
  Pclass="3rd",
  Parch=2, 
  Adult_male=0,
  SibSp=0)

# Sue odds with and without Leonardo
Sue_odds <-  exp(predict(finalmodel,Sue))
Sue_odds_Leo <-  exp(predict(finalmodel, Sue_with_Leo))

# Sue probability with and without Leonardo
Sue_prob <- Sue_odds / (1 + Sue_odds)
Sue_prob_Leo <-  Sue_odds_Leo / (1+ Sue_odds_Leo)

# Result
summary(Sue_prob_Leo)
summary(Sue_prob)
```


```{r}
predict(finalmodel,newdata=Kate)
pR2(finalmodel)
# -2LL, deviance	
pR2(finalmodel)["llh"] * -2	
```

Mutate to predict survival
```{r}
titanic_na = titanic_na %>% 
  mutate(pred_finalmodel = predict(finalmodel)) %>%
  mutate(pred_finalmodel = case_when(pred_finalmodel <= 0 ~ "no",
                                 pred_finalmodel > 0 ~ "yes"))
```

Coding the correct quesses
```{r}
titanic_na = titanic_na %>%	
  mutate(correct_prediction = case_when(pred_finalmodel == Survived ~ "correct",	
                                        pred_finalmodel != Survived ~ "incorrect"))	
```

Percentage of survivel overall for the model
```{r}
titanic_na %>%	
  group_by(correct_prediction) %>%	
  summarise(count = n()) %>%	
  mutate(freq = count / sum(count))	
```

Survived and predicted values displayed in a crosstab
```{r}
titanic_na %>% 	
  group_by(Survived, pred_finalmodel) %>% 	
  summarize(n = n()) %>% 	
  spread(Survived, n)	
```

Percentage of correctly predicted survival
```{r}
titanic_na %>%	
  filter(Survived == "yes") %>% 	
  group_by(correct_prediction) %>%	
  summarise(count = n()) %>%	
  mutate(freq = count / sum(count))	
```

Percentage of correctly predicted for those who did not survive
```{r}
titanic_na %>%	
  filter(Survived == "no") %>% 	
  group_by(correct_prediction) %>%	
  summarise(count = n()) %>%	
  mutate(freq = count / sum(count))	
```

As i am satisfied with my model I chose not to fine tune it. 
