---
title: "Linear_mixed_models-lab"
author: "Majbritt Lindholt Stenbøg"

packages used:
```{r}
library(lme4) 
library(r2glmm)
library(MuMIn) 
library(lmerTest) 
library(tidyverse) 	
library(dplyr)
library(gridExtra) 
```

This is a function to extract standardized beta coefficients from linear mixed models:
This function was adapted from: https://stackoverflow.com/questions/25142901/standardized-coefficients-for-lmer-model:
```{r}
stdCoef.merMod <- function(object) {	
  sdy <- sd(getME(object,"y"))	
  sdx <- apply(getME(object,"X"), 2, sd)	
  sc <- fixef(object)*sdx/sdy	
  se.fixef <- coef(summary(object))[,"Std. Error"]	
  se <- se.fixef*sdx/sdy	
  return(data.frame(stdcoef=sc, stdse=se))	
}
```

At first i load the dataset:
```{r}
surgeryDataA <- read.table("surgery_data_A.csv",header=TRUE,sep=",")

view(surgeryDataA)
```

Checking the data:
```{r}
unique(surgeryDataA$pain)

unique(surgeryDataA$pain_cat)

unique(surgeryDataA$hospital)

# mindfulness
surgeryDataA %>%
  ggplot() +
  aes(x = mindfulness) +
  geom_bar()

# Pain
surgeryDataA %>%
  ggplot() +
  aes(x = pain) +
  geom_bar()

# STAI_trait
surgeryDataA %>%
  ggplot() +
  aes(x = STAI_trait) +
  geom_bar()

# Age
surgeryDataA %>%
  ggplot() +
  aes(x = age) +
  geom_bar()

# Cortisol serum
surgeryDataA %>%
  ggplot() +
  aes(x = cortisol_serum) +
  geom_histogram()

# Sex
surgeryDataA %>%
  ggplot() +
  aes(x = sex) +
  geom_bar() +
  scale_x_discrete(labels=c("Male","Female"))

# Household income for dataset A
fig1 <- surgeryDataA %>%
  ggplot() +
  aes(y = household_income) + 
  geom_boxplot() +
  ylim(c(-8000,135000)) +
  ggtitle("Household income, Dataset A")

# Household income for dataset B
fig2 <- surgeryDataB %>%
  ggplot() +
  aes(y = household_income) + 
  geom_boxplot() +
  ylim(c(11000,145000)) +
  ggtitle("Household income, Dataset B")

grid.arrange(fig1,fig2,nrow=1)

table(surgeryDataA$mindfulness)
table(surgeryDataA$pain_cat)
table(surgeryDataA$STAI_trait)
summary(surgeryDataA)

head(surgeryDataA)
```

I want to recode sex for easier interpretation:
```{r}
surgeryDataA <- surgeryDataA %>%
  mutate(sex = factor(recode(sex,"male"="0", "female"="1","woman"="1")))

view(surgeryDataA)

unique(surgeryDataA$sex)

table(surgeryDataA$sex)
```

In the following i check for clustering: 
```{r}
# Checking clustering pain + pain_cat
surgeryDataA %>%
  ggplot() + 
  aes(y = pain, x = pain_cat) + 
  geom_point(aes(color = hospital),size = 4) + 
  geom_smooth(method = "lm", se = F)

# Checking clustering pain + mindfulness
surgeryDataA %>%
  ggplot() + 
  aes(y = pain, x = mindfulness) + 
  geom_point(aes(color = hospital), size = 4) + 
  geom_smooth(method = "lm", se = F, 
              aes(color = hospital))
```

I examine the data with a linear model first:
```{r}
fixed_effect_p <- lm(pain  ~ age + sex + STAI_trait + pain_cat + mindfulness + cortisol_serum, 
      data= surgeryDataA)
      
summary(fixed_effect_p)
```

Now i want to built a random intercept model with hospital-ID:
```{r}
random_intercept = lmer(pain ~ age + sex + STAI_trait + pain_cat + mindfulness + cortisol_serum + (1 | hospital), 
      data = surgeryDataA)

summary(random_intercept)

# confidence intervals
confint(random_intercept, level = 0.95)

# Standardized beta for each predictor:	
stdCoef.merMod(random_intercept)	

```

Marginal and conditional R2:
```{r}
# marginal R squared with confidence intervals - 	
r2beta(random_intercept, 
       method = "nsj", 
       data = surgeryDataA)	

# model had an explanability of 38.5%
# the predictors that best explain pain is:
# cortisol_serum

# marginal and conditional R squared values	- 
r.squaredGLMM(random_intercept)	
```

Load datafile B:
```{r}
surgeryDataB <- read.table("surgery_data_B.csv",header=TRUE,sep=",")

view(surgeryDataB)
```

We should always check the data before we work with it:
```{r}
unique(surgeryDataB$pain)
unique(surgeryDataB$pain_cat)
unique(surgeryDataB$hospital)

# mindfulness
surgeryDataB %>%
  ggplot() +
  aes(x = mindfulness) +
  geom_bar()

# pain
surgeryDataB %>%
  ggplot() +
  aes(x = pain) +
  geom_bar()

# STAI_trait
surgeryDataB %>%
  ggplot() +
  aes(x = STAI_trait) +
  geom_bar()

# Age
surgeryDataB %>%
  ggplot() +
  aes(x = age) +
  geom_bar()

# cortisol serum
surgeryDataB %>%
  ggplot() +
  aes(x = cortisol_serum) +
  geom_histogram()

# Sex
surgeryDataB %>%
  ggplot() +
  aes(x = sex) +
  geom_bar() +
  scale_x_discrete(labels=c("Male","Female"))

#This also has an outlier like Dataset A. This is just a very high household income than a low one.
surgeryDataB %>%
  ggplot() +
  aes(y = household_income) + 
  geom_boxplot()

summary(surgeryDataB)

head(surgeryDataB)
unique(surgeryDataB$sex)
```

I want to recode sex again so the two datasets fits:
```{r}
surgeryDataB <- surgeryDataB %>%
  mutate(sex = factor(recode(sex,"male"="0", "female"="1")))

unique(surgeryDataB$sex)

table(surgeryDataB$sex)
```

first look at the predicted values for surgery data A:
```{r}
surgeryDataA <-  surgeryDataA %>%
  mutate(predict_A = predict(random_intercept))

# Plot the predicted
surgeryDataA %>%
  ggplot() + 
  aes(x = pain, y = pain_cat, group = hospital) +
  geom_point(aes(color = hospital), size=1.3) +
  geom_line(color = "blue", aes(x=predict_A, y=pain_cat)) + 
  facet_wrap(~hospital,ncol=2)

```

use the model coefficients on data file A to predict data file B:
```{r}
predict_B <- predict(random_intercept, 
                     newdata=surgeryDataB, 
                     allow.new.levels = TRUE)

# This is the predicted pain for data file B
predict_B

# Add the value to the dataset
surgeryDataB_wPredict <- cbind(surgeryDataB, predict_B)

# Final plot for predict pain with model from data file A in data file B 
ggplot(data=surgeryDataA, 
      aes(x = predict_A, y=pain)) + 
  geom_point(aes(color="Predicted A"), fill="green") + 
  geom_point(data=surgeryDataB_wPredict,
      aes(x=predict_B, y=pain, color="Predicted B"), fill="red") + 
  scale_color_manual("Predicted values", 
      limits=c("Predicted A", "Predicted B"), 
      values = c("green","red")) +
  theme(
      legend.background = element_rect(fill = "white", colour = "black"),
      plot.title = element_text(
      size = rel(1.2), lineheight = .9,
      family = "Times New Roman", face = "bold", colour = "black")) + 
  ggtitle("Pain Predicted for Data file B") +
  labs(x="Predicted",y="Pain")

#The two plots side by side
ggplot(surgeryDataA,
      aes(predict_A,pain)) + 
      geom_point() 
ggplot(surgeryDataB_wPredict,
      aes(predict_B,pain)) + 
      geom_point()
```

Variance explained:
```{r}
#create null model for TSS
null_model <- lm(pain ~1, data=surgeryDataB)

# Create the same model for data file B - use your model where you predicted for datafile B:
predict_B

# RSS
RSS <- sum((surgeryDataB$pain-predict_B)^2)

# TSS
TSS <- sum((surgeryDataB$pain-predict(null_model))^2)

# Variance explained
R2 <- 1-(RSS/TSS)
R2

#it is clsoer to the marginal R2
# since we built the model only with the A data set we don't have any information about the hospital (random effect) in the B data set so we can only explain variance with the fixed effects  -> thus it is closer to the marginal R2 of the A data
```

New model with only cortisol_serum:
```{r}
# based on the R^2 test of the random_intercept model it showed that cortisol_serum were the strongest predictor of pain. 

#random intercept model
r_intercept_final <- lmer(pain ~ cortisol_serum + (1 | hospital), 
                         data = surgeryDataA)

summary(r_intercept_final)

# random slope model
r_slope_final <- lmer(pain ~ cortisol_serum + ( cortisol_serum | hospital), 
                     data = surgeryDataA)

summary(r_slope_final)
summary(r_intercept_final)
#getting an error of singular fit
```

Now visualize the fitted regression lines for each hospital separately.:
vizualising the models:
```{r}
# saving the predictions of the models into a variable.
surgeryDataA <-  surgeryDataA %>%
mutate(pred_int_final = predict(r_intercept_final), pred_slope_final = predict(r_slope_final))

# Model for the random intercept
LMM_1 <- surgeryDataA %>%
  ggplot() + 
  aes(y = pain, x = cortisol_serum, group = hospital) +
  geom_point(aes(color = hospital), size = 1.5) +
  geom_line(color = "red",
  aes(y = pred_int_final, x = cortisol_serum)) + 
   ggtitle("Random intercept model") +
  facet_wrap(~hospital, ncol = 2)

# Model for the random slope 
LMM_2 <- surgeryDataA %>%
  ggplot() + 
  aes(y = pain, x = cortisol_serum, group = hospital) +
  geom_point(aes(color = hospital), size = 1.5) + 
  geom_line(color = "red",
  aes(y = pred_slope_final, x = cortisol_serum)) + 
   ggtitle("Random slope model") +
  facet_wrap(~hospital, ncol = 2)

grid.arrange(LMM_1,LMM_2,nrow=1)

anova(r_intercept_final,r_slope_final)

```



