---
title: "lab assignment 3"
author: "Majbritt Lindholt Stenbøg"
---
Packages used:
```{r}

library(psych) 
library(car) 
library(GGally) 	
library(ggcorrplot)
library(GPArotation) 	
library(MVN) 
library(ICS) 
library(FactoMineR) 
library(factoextra) 
library(paran) 
library(gridExtra)

# Custom function
fviz_loadnings_with_cor <- function(mod, axes = 1, loadings_above = 0.4){	
  require(factoextra)	
  require(dplyr)	
  require(ggplot2)	
	
	
	
if(!is.na(as.character(mod$call$call)[1])){	
  if(as.character(mod$call$call)[1] == "PCA"){	
  contrib_and_cov = as.data.frame(rbind(mod[["var"]][["contrib"]], mod[["var"]][["cor"]]))	
	
vars = rownames(mod[["var"]][["contrib"]])	
attribute_type = rep(c("contribution","correlation"), each = length(vars))	
contrib_and_cov = cbind(contrib_and_cov, attribute_type)	
contrib_and_cov	
	
plot_data = cbind(as.data.frame(cbind(contrib_and_cov[contrib_and_cov[,"attribute_type"] == "contribution",axes], contrib_and_cov[contrib_and_cov[,"attribute_type"] == "correlation",axes])), vars)	
names(plot_data) = c("contribution", "correlation", "vars")	
	
plot_data = plot_data %>% 	
  mutate(correlation = round(correlation, 2))	
	
plot = plot_data %>% 	
  ggplot() +	
  aes(x = reorder(vars, contribution), y = contribution, gradient = correlation, label = correlation)+	
  geom_col(aes(fill = correlation)) +	
  geom_hline(yintercept = mean(plot_data$contribution), col = "red", lty = "dashed") + scale_fill_gradient2() +	
  xlab("variable") +	
  coord_flip() +	
  geom_label(color = "black", fontface = "bold", position = position_dodge(0.5))	
	
	
}	
} else if(!is.na(as.character(mod$Call)[1])){	
  	
  if(as.character(mod$Call)[1] == "fa"){	
    loadings_table = mod$loadings %>% 	
      matrix(ncol = ncol(mod$loadings)) %>% 	
      as_tibble() %>% 	
      mutate(variable = mod$loadings %>% rownames()) %>% 	
      gather(factor, loading, -variable) %>% 	
      mutate(sign = if_else(loading >= 0, "positive", "negative"))	
  	
  if(!is.null(loadings_above)){	
    loadings_table[abs(loadings_table[,"loading"]) < loadings_above,"loading"] = NA	
    loadings_table = loadings_table[!is.na(loadings_table[,"loading"]),]	
  }	
  	
  if(!is.null(axes)){	
  	
  loadings_table = loadings_table %>% 	
     filter(factor == paste0("V",axes))	
  }	
  	
  	
  plot = loadings_table %>% 	
      ggplot() +	
      aes(y = loading %>% abs(), x = reorder(variable, abs(loading)), fill = loading, label =       round(loading, 2)) +	
      geom_col(position = "dodge") +	
      scale_fill_gradient2() +	
      coord_flip() +	
      geom_label(color = "black", fill = "white", fontface = "bold", position = position_dodge(0.5)) +	
      facet_wrap(~factor) +	
      labs(y = "Loading strength", x = "Variable")	
  }	
}	
	
	
	
	
	
	
return(plot)	
	
}	
```

Load dataset: 
```{r}
animalrights <- read.csv("animalrights.csv",header=TRUE,sep=",")

view(animalrights)
```

Explore it:
```{r}
unique(animalrights$liberal)
unique(animalrights_na$ar1)
unique(animalrights$ar2)
unique(animalrights$ar3)
unique(animalrights$ar4)
unique(animalrights_na$ar5)
unique(animalrights_na$ar7)
describe(animalrights)
```

Important to note about my dataset: 
no "5" answer: ar3, ar4, 
NA's: ar1, ar2, ar10, ar14, ar18, ar19, ar22, liberal
So important to explore the data and consider treating this data. I would like to exclude NA's because they will be excluded by the analysis anyways:
```{r}
animalrights_na <- drop_na(animalrights)

view(animalrights_na)
```

Checking data:
```{r}
# ar3
animalrights_na %>%
  ggplot() +
  aes(x = ar3) +
  geom_histogram()

# Liberal
animalrights_na %>%
  ggplot() +
  aes(x = liberal) +
  geom_histogram()

# Liberal
animalrights_na %>%
  ggplot() +
  aes(x = liberal) +
  geom_boxplot()

# party, most are coded as none so not helpful
animalrights_na %>%
  ggplot() +
  aes(x = party) +
  geom_histogram()
table(animalrights_na$party)

# recode sex because it's not a factor
animalrights_na %>%
  ggplot() +
  aes(x = sex) +
  geom_bar()
table(animalrights_na$sex)

#recoding sex:
animalrights_na$sex <- factor(animalrights_na$sex)

summary(animalrights_na)

str(animalrights_na)

animalrights_na %>% describe()
```

A linear model and VIF test:
```{r}
# We can make a linear model 
lm_allQ <- lm(liberal ~ ar1+ ar2+ ar3+ ar4+ ar5+ ar6+ ar7+ ar8+ ar9+ ar10+ ar11+ ar12+ ar13+ ar14+ ar15+ ar16+ ar17+ ar18+ ar19+ ar20+ ar21+ ar22+ ar23+ ar24+ ar25+ ar26+ ar27+ ar28, data = animalrights_na)

summary(lm_allQ)

# VIF test
vif(lm_allQ)

# most are under 3 which shows no multicollinearity, except for ar5, ar13, which is high, and ar6 and ar15 is close.
```

Vizualising the items:
```{r}
ARQ_only <- animalrights_na %>%
  select(ar1:ar28)

cor_ARQ <- ARQ_only %>%
    cor()

ggcorr(cor_ARQ)

ggcorrplot(cor(ARQ_only), p.mat = cor_pmat(ARQ_only), hc.order = TRUE, type="lower")
```

correlation matrix:
```{r}
ARQ_cor <- cor(animalrights_na[1:28])
ARQ_cor
```
Bartlett sphericity test:
in our case, the number of observations is 149 and observed variables is 28 which is = 5,33. 
it is still more than 5 but close so we can try to run it:
```{r}
Barlett_factorability <- cortest.bartlett(ARQ_cor)
Barlett_factorability
```
As we can see it is very significant which means that there is correlation in our dataset.

KMO:
```{r}
KMO(ARQ_cor)
```

multivariate normal distribution tests:
```{r}
# test for multiariate normality
mvn_result <- mvn(animalrights_na[, 1:28], mvnTest = "hz")
mvn_result$multivariateNormality
```

```{r}
# multivariate normality test  based on kurtosis
mvnorm.kur.test(na.omit(animalrights_na[, 1:28]))
```

```{r}
# Multivariate normality test based on skewness
mvnorm.skew.test(na.omit(animalrights_na[, 1:28]))
```

First model test:
```{r}
# i will use the fa() finction to make my first EFA model. here we will set the fm = parameter to "pa" since we want to use the paf method because of the multivariate normality violation. with the nfactors = we can specify how many factors we want to extract. we will try first with 5. 

AR_EFAmod1 <- fa(ARQ_cor, nfactors = 5, fm = "pa")
AR_EFAmod1
AR_EFAmod1_c <- as.data.frame(sort(AR_EFAmod1$communality, decreasing = TRUE))
AR_EFAmod1_c
mean(AR_EFAmod1$communality)
```

**PCA**
```{r}
PCA_model <- pca(ARQ_only)
PCA_model

#make a new PCA model
which(names(animalrights_na) == "sex")
which(names(animalrights_na) == "party")
which(names(animalrights_na) == "liberal")

PCA_model2 <- PCA(animalrights_na, quanti.sup = c(30,31), quali.sup = 29)

summary(PCA_model2)
```

I need to figure out how many factors i want to extract from the data:
```{r}
# Screetest

fviz_screeplot(PCA_model2, addlabels = TRUE, ylim = c(0, 85))	

#Kaiser guttman rule
get_eigenvalue(PCA_model2)

# Parallel test
PCA_parallel = paran(ARQ_only, graph = TRUE)
PCA_parallel$Retained

```

```{r}
# parallel test
fa.parallel(ARQ_cor, n.obs = nrow(animalrights_na), fa = "fa", fm = "pa")
```

VSS and MAP:
```{r}
# VSS and MAP test
nfactors(ARQ_cor, n.obs = nrow(animalrights_na))

VSS_test <-  vss(ARQ_cor, n.obs =nrow(animalrights_na))

VSS.plot(VSS_test)
```

EFA model with 2 factors:
```{r}
AR_EFAmod2 <-  fa(ARQ_cor, nfactors = 2, fm = "pa")
AR_EFAmod2
AR_EFAmod2_c <- as.data.frame(sort(AR_EFAmod2$communality, decreasing = TRUE))
AR_EFAmod2_c
mean(AR_EFAmod2$communality)
```

Factor rotation:
```{r}
AR_EFAmod2$rotation

# Varimax
AR_EFAmod2_varimax <- fa(ARQ_cor, nfactors = 2, fm = "pa", rotate = "varimax")
AR_EFAmod2_varimax 
# Promax
AR_EFAmod2_promax <- fa(ARQ_cor, nfactors = 2, fm = "pa", rotate = "promax")
AR_EFAmod2_promax
# Oblimin
AR_EFAmod2_oblimin <- fa(ARQ_cor, nfactors = 2, n.obs =nrow(animalrights_na), fm = "pa", rotate = "oblimin")
AR_EFAmod2_oblimin

AR_EFAmod2_promax$communality
```

Diagram function:
```{r}
fa.diagram(AR_EFAmod2_varimax )

fviz_loadnings_with_cor(AR_EFAmod2_oblimin, axes = 1, loadings_above = 0.4)
```

New model building. i extracted ar16, because it didn't have any high correlation in any of the factors, further, by looking at the questions, it might be a causal relationship or be a result of confounding, as it was a religious question, which didn't fit into the themes of the other questions:
```{r}
cormod3 <- cor(animalrights_na[, - c(16,29,30,31)])
AR_EFAmod3_varimax <- fa(cormod3, nfactors = 2, fm = "pa", rotate = "varimax")
AR_EFAmod3_varimax 

AR_EFAmod3_promax <- fa(cormod3, nfactors = 2, fm = "pa", rotate = "promax")
AR_EFAmod3_promax

mean(AR_EFAmod3_promax$communality)

print(AR_EFAmod3_promax$loadings, cutoff = 0.3)
fa.diagram(AR_EFAmod3_varimax )
```

I would like to remove ar11 as it seems to have a very low loading
```{r}
cormod4 <- cor(animalrights_na[, - c(11,16,29,30,31)])

AR_EFAmod4_promax <- fa(cormod4, nfactors = 2, fm = "pa", rotate = "promax")
AR_EFAmod4_promax

mean(AR_EFAmod4_promax$communality)
AR_EFAmod4_promax$communality

print(AR_EFAmod4_promax$loadings, cutoff = 0.3)

AR_EFAmod4_c <- as.data.frame(sort(AR_EFAmod4_promax$communality, decreasing = TRUE))
AR_EFAmod4_c
```
After a few tries, i change it to three factors as it covers the theory better, and improves communality:
```{r}
cormod3.2 <- cor(animalrights_na[, - c(16,29,30,31)])
AR_EFAmod3_varimax <- fa(cormod3.2, nfactors = 3, fm = "pa", rotate = "varimax")
AR_EFAmod3_varimax 

AR_EFAmod3_promax <- fa(cormod3.2, nfactors = 3, fm = "pa", rotate = "promax")
AR_EFAmod3_promax

mean(AR_EFAmod3$communality)
print(AR_EFAmod3_promax$loadings, cutoff = 0.3)

fa.diagram(AR_EFAmod3_promax)
?fa.diagram
AR_EFAmod3_promax$communality
```
I would like to remove ar12 as it seems to be in between 2 factors:
```{r}
cormod4.2 <- cor(animalrights_na[, - c(12,16,29,30,31)])

AR_EFAmod4_promax <- fa(cormod4.2, nfactors = 3, fm = "pa", rotate = "promax")
AR_EFAmod4_promax

mean(AR_EFAmod4_promax$communality)

print(AR_EFAmod4_promax$loadings, cutoff = 0.3)

AR_EFAmod4_c <- as.data.frame(sort(AR_EFAmod4_promax$communality, decreasing = TRUE))
AR_EFAmod4_c

fa.diagram(AR_EFAmod4_promax)

```
ar8 has a very low communality (0.15), which is why i want to remove it: 
```{r}
cormod5 <- cor(animalrights_na[, - c(8,12,16,29,30,31)])
AR_EFAmod5 <- fa(cormod5, nfactors = 3, fm = "pa")
AR_EFAmod5_promax <- fa(cormod5, nfactors = 3, fm = "pa", rotate = "promax")
AR_EFAmod5_promax

mean(AR_EFAmod5_promax$communality)

print(AR_EFAmod5_promax$loadings, cutoff = 0.3)

AR_EFAmod5_c <- as.data.frame(sort(AR_EFAmod5_promax$communality, decreasing = TRUE))
AR_EFAmod5_c

AR_EFAmod5_varimax <- fa(ARQ_cor[, - c(12,15,16)], nfactors = 2, fm = "pa", rotate = "varimax")
AR_EFAmod5_varimax
print(AR_EFAmod5_varimax$loadings, cutoff = 0.3)
# doesn't seem as if varimax works for my analysis. I will continue using promax. I allow some overlapping as some of the questions are close to each other, and all concern animal rights.
fa.diagram(AR_EFAmod5_promax )

fviz_loadnings_with_cor(AR_EFAmod5_promax, axes = 1, loadings_above = 0.4)
fviz_loadnings_with_cor(AR_EFAmod5_promax, axes = 2, loadings_above = 0.4)
```
Question ar3 is so low, that it is not even loading at the 0.3 mark of main loadings. i remove it. 
```{r}
cormod6 <- cor(animalrights_na[, - c(3,8,12,16,29,30,31)])

AR_EFAmod6_promax <- fa(cormod6, nfactors = 3, fm = "pa", rotate = "promax")
AR_EFAmod6_promax

AR_EFAmod6_oblimin <- fa(cormod6, nfactors = 3, fm = "pa", rotate = "oblimin")
AR_EFAmod6_oblimin

mean(AR_EFAmod6_promax$communality)

print(AR_EFAmod6_promax$loadings, cutoff = 0.3)
print(AR_EFAmod6_oblimin$loadings, cutoff = 0.3)

AR_EFAmod6_c <- as.data.frame(sort(AR_EFAmod6_promax$communality, decreasing = TRUE))
AR_EFAmod6_c

fviz_loadnings_with_cor(AR_EFAmod6_promax, axes = 2, loadings_above = 0.4)
```
now ar14 is super low and will be removed as well. it also has the lowest communality:
```{r}
cormod7 <- cor(animalrights_na[, - c(3,8,12,14,16,29,30,31)])

AR_EFAmod7_promax <- fa(cormod7, nfactors = 3, fm = "pa", rotate = "promax")
AR_EFAmod7_promax

mean(AR_EFAmod7_promax$communality)

print(AR_EFAmod7_promax$loadings, cutoff = 0.3)
fa.diagram(AR_EFAmod7_promax)
```
Further, i want to remove ar18 because it is giving opposite results, and the main loadings are low in both, creating cross loadings:
```{r}
cormod8 <- cor(animalrights_na[, - c(3,8,12,14,16,18,29,30,31)])

AR_EFAmod8_promax <- fa(cormod8, nfactors = 3, fm = "pa", rotate = "promax")
AR_EFAmod8_promax

mean(AR_EFAmod8_promax$communality)
AR_EFAmod8_promax$communality

print(AR_EFAmod8_promax$loadings, cutoff = 0.3)
fa.diagram(AR_EFAmod7_promax)
```
as ar1 seems to be the one with lowest loadings and low communality, i will remove it next:
```{r}
cormod9 <- cor(animalrights_na[, - c(1,3,8,12,14,16,18,29,30,31)])

AR_EFAmod9_promax <- fa(cormod9, nfactors = 3, fm = "pa", rotate = "promax")
AR_EFAmod9_promax

mean(AR_EFAmod9_promax$communality)

print(AR_EFAmod9_promax$loadings, cutoff = 0.3)
fa.diagram(AR_EFAmod9_promax)

fviz_loadnings_with_cor(AR_EFAmod9_promax, axes = 1, loadings_above = 0.4)
fviz_loadnings_with_cor(AR_EFAmod9_promax, axes = 2, loadings_above = 0.4)
```
ar11 is removed as it has no function in factor 2:
```{r}
cormod10 <- cor(animalrights_na[, - c(1,3,8,11,12,14,16,18,29,30,31)])

AR_EFAmod10_promax <- fa(cormod10, nfactors = 3, fm = "pa", rotate = "promax")
AR_EFAmod10_promax

mean(AR_EFAmod10_promax$communality)
AR_EFAmod10_promax$communality

print(AR_EFAmod10_promax$loadings, cutoff = 0.3)
fa.diagram(AR_EFAmod10_promax)

fviz_loadnings_with_cor(AR_EFAmod10_promax, axes = 1, loadings_above = 0.4) + labs(title= "Aganst Using Animals in Research")
fviz_loadnings_with_cor(AR_EFAmod10_promax, axes = 2, loadings_above = 0.4) + labs(title= "Pro Vegan Attitudes")
fviz_loadnings_with_cor(AR_EFAmod10_promax, axes = 3, loadings_above = 0.4) + labs(title= "Traditional Attitudes to Animals")
```
add model to dataframe:
```{r}
Factors <- factor.scores(animalrights_na[,-c(1,3,8,11,12,14,16,18,29,30,31)],AR_EFAmod10_promax)$scores
colnames(Factors) <-  c("research", "pro_vegan","traditional")

view(Factors)

final_data <- cbind(animalrights_na[,-c(1,3,8,11,12,14,16,18)],Factors)

view(final_data)
```

now i want to make a linear model predicting how conservative or liberal the participants are:
```{r}
plot(final_data$liberal,final_data$research)

plot(final_data$liberal, final_data$animal_killing)

predict_liberal <-  lm(liberal ~ research + pro_vegan + traditional, data= final_data)

summary(predict_liberal)

vif(predict_liberal) # ikke multicollinearitet
```
 
