---
title: "lab assignment 3"
author: "Majbritt Lindholt Stenbøg"
---
Packages used:
```{r}

library(psych) 
library(car) 
library(GGally) 	
library(ggcorrplot)
library(GPArotation) 	
library(MVN) 
library(ICS) 
library(FactoMineR) 
library(factoextra) 
library(paran) 
library(gridExtra)

# Custom function
fviz_loadnings_with_cor <- function(mod, axes = 1, loadings_above = 0.4){	
  require(factoextra)	
  require(dplyr)	
  require(ggplot2)	
	
	
	
if(!is.na(as.character(mod$call$call)[1])){	
  if(as.character(mod$call$call)[1] == "PCA"){	
  contrib_and_cov = as.data.frame(rbind(mod[["var"]][["contrib"]], mod[["var"]][["cor"]]))	
	
vars = rownames(mod[["var"]][["contrib"]])	
attribute_type = rep(c("contribution","correlation"), each = length(vars))	
contrib_and_cov = cbind(contrib_and_cov, attribute_type)	
contrib_and_cov	
	
plot_data = cbind(as.data.frame(cbind(contrib_and_cov[contrib_and_cov[,"attribute_type"] == "contribution",axes], contrib_and_cov[contrib_and_cov[,"attribute_type"] == "correlation",axes])), vars)	
names(plot_data) = c("contribution", "correlation", "vars")	
	
plot_data = plot_data %>% 	
  mutate(correlation = round(correlation, 2))	
	
plot = plot_data %>% 	
  ggplot() +	
  aes(x = reorder(vars, contribution), y = contribution, gradient = correlation, label = correlation)+	
  geom_col(aes(fill = correlation)) +	
  geom_hline(yintercept = mean(plot_data$contribution), col = "red", lty = "dashed") + scale_fill_gradient2() +	
  xlab("variable") +	
  coord_flip() +	
  geom_label(color = "black", fontface = "bold", position = position_dodge(0.5))	
	
	
}	
} else if(!is.na(as.character(mod$Call)[1])){	
  	
  if(as.character(mod$Call)[1] == "fa"){	
    loadings_table = mod$loadings %>% 	
      matrix(ncol = ncol(mod$loadings)) %>% 	
      as_tibble() %>% 	
      mutate(variable = mod$loadings %>% rownames()) %>% 	
      gather(factor, loading, -variable) %>% 	
      mutate(sign = if_else(loading >= 0, "positive", "negative"))	
  	
  if(!is.null(loadings_above)){	
    loadings_table[abs(loadings_table[,"loading"]) < loadings_above,"loading"] = NA	
    loadings_table = loadings_table[!is.na(loadings_table[,"loading"]),]	
  }	
  	
  if(!is.null(axes)){	
  	
  loadings_table = loadings_table %>% 	
     filter(factor == paste0("V",axes))	
  }	
  	
  	
  plot = loadings_table %>% 	
      ggplot() +	
      aes(y = loading %>% abs(), x = reorder(variable, abs(loading)), fill = loading, label =       round(loading, 2)) +	
      geom_col(position = "dodge") +	
      scale_fill_gradient2() +	
      coord_flip() +	
      geom_label(color = "black", fill = "white", fontface = "bold", position = position_dodge(0.5)) +	
      facet_wrap(~factor) +	
      labs(y = "Loading strength", x = "Variable")	
  }	
}	
	
	
	
	
	
	
return(plot)	
	
}	
```

Load dataset: 
```{r}
animalrights <- read.csv("animalrights.csv",header=TRUE,sep=",")

view(animalrights)
```

Explore it:
```{r}
unique(animalrights$liberal)
unique(animalrights_na$ar1)
unique(animalrights$ar2)
unique(animalrights$ar3)
unique(animalrights$ar4)
unique(animalrights_na$ar5)
unique(animalrights_na$ar7)
describe(animalrights)
```

Important to note about my dataset: 
no "5" answer: ar3, ar4, 
NA's: ar1, ar2, ar10, ar14, ar18, ar19, ar22, liberal
So important to explore the data and consider treating this data. I would like to exclude NA's because they will be excluded by the analysis anyways:
```{r}
animalrights_na <- drop_na(animalrights)

view(animalrights_na)
```

Checking data:
```{r}
# ar3
animalrights_na %>%
  ggplot() +
  aes(x = ar3) +
  geom_histogram()

# Liberal
animalrights_na %>%
  ggplot() +
  aes(x = liberal) +
  geom_histogram()

# Liberal
animalrights_na %>%
  ggplot() +
  aes(x = liberal) +
  geom_boxplot()

# party, most are coded as none so not helpful
animalrights_na %>%
  ggplot() +
  aes(x = party) +
  geom_histogram()
table(animalrights_na$party)

# recode sex because it's not a factor
animalrights_na %>%
  ggplot() +
  aes(x = sex) +
  geom_bar()
table(animalrights_na$sex)

#recoding sex:
animalrights_na$sex <- factor(animalrights_na$sex)

summary(animalrights_na)

str(animalrights_na)

animalrights_na %>% describe()
```

A linear model and VIF test:
```{r}
# We can make a linear model 
lm_allQ <- lm(liberal ~ ar1+ ar2+ ar3+ ar4+ ar5+ ar6+ ar7+ ar8+ ar9+ ar10+ ar11+ ar12+ ar13+ ar14+ ar15+ ar16+ ar17+ ar18+ ar19+ ar20+ ar21+ ar22+ ar23+ ar24+ ar25+ ar26+ ar27+ ar28, data = animalrights_na)

summary(lm_allQ)

# VIF test
vif(lm_allQ)

# most are under 3 which shows no multicollinearity, except for ar5, ar13, which is high, and ar6 and ar15 is close.
```

Vizualising the items:
```{r}
ARQ_only <- animalrights_na %>%
  select(ar1:ar28)

cor_ARQ <- ARQ_only %>%
    cor()

ggcorr(cor_ARQ)

ggcorrplot(cor(ARQ_only), p.mat = cor_pmat(ARQ_only), hc.order = TRUE, type="lower")
```

correlation matrix:
```{r}
ARQ_cor <- cor(animalrights_na[1:28])
ARQ_cor
```
Bartlett sphericity test:
in our case, the number of observations is 149 and observed variables is 28 which is = 5,33. 
it is still more than 5 but close so we can try to run it:
```{r}
Barlett_factorability <- cortest.bartlett(ARQ_cor)
Barlett_factorability
```
As we can see it is very significant which means that there is correlation in our dataset.

KMO:
```{r}
KMO(ARQ_cor)
```

multivariate normal distribution tests:
```{r}
# test for multiariate normality
mvn_result <- mvn(animalrights_na[, 1:28], mvnTest = "hz")
mvn_result$multivariateNormality
```

```{r}
# multivariate normality test  based on kurtosis
mvnorm.kur.test(na.omit(animalrights_na[, 1:28]))
```

```{r}
# Multivariate normality test based on skewness
mvnorm.skew.test(na.omit(animalrights_na[, 1:28]))
```

First model test:
```{r}
# i will use the fa() finction to make my first EFA model. here we will set the fm = parameter to "pa" since we want to use the paf method because of the multivariate normality violation. with the nfactors = we can specify how many factors we want to extract. we will try first with 5. 

AR_EFAmod1 <- fa(ARQ_cor, nfactors = 5, fm = "pa")
AR_EFAmod1
AR_EFAmod1_c <- as.data.frame(sort(AR_EFAmod1$communality, decreasing = TRUE))
AR_EFAmod1_c
mean(AR_EFAmod1$communality)
```

**PCA**
```{r}
PCA_model <- pca(ARQ_only)
PCA_model

#make a new PCA model
which(names(animalrights_na) == "sex")
which(names(animalrights_na) == "party")
which(names(animalrights_na) == "liberal")

PCA_model2 <- PCA(animalrights_na, quanti.sup = c(30,31), quali.sup = 29)

summary(PCA_model2)
```

I need to figure out how many factors i want to extract from the data:
```{r}
# Screetest

fviz_screeplot(PCA_model2, addlabels = TRUE, ylim = c(0, 85))	

#Kaiser guttman rule
get_eigenvalue(PCA_model2)

# Parallel test
PCA_parallel = paran(ARQ_only, graph = TRUE)
PCA_parallel$Retained

```

```{r}
# parallel test
fa.parallel(ARQ_cor, n.obs = nrow(animalrights_na), fa = "fa", fm = "pa")
```

VSS and MAP:
```{r}
# VSS and MAP test
nfactors(ARQ_cor, n.obs = nrow(animalrights_na))

VSS_test <-  vss(ARQ_cor, n.obs =nrow(animalrights_na))

VSS.plot(VSS_test)
```

EFA model with 2 factors:
```{r}
AR_EFAmod2 <-  fa(ARQ_cor, nfactors = 2, fm = "pa")
AR_EFAmod2
AR_EFAmod2_c <- as.data.frame(sort(AR_EFAmod2$communality, decreasing = TRUE))
AR_EFAmod2_c
mean(AR_EFAmod2$communality)
```

Factor rotation:
```{r}
AR_EFAmod2$rotation

# Varimax
AR_EFAmod2_varimax <- fa(ARQ_cor, nfactors = 2, fm = "pa", rotate = "varimax")
AR_EFAmod2_varimax 
# Promax
AR_EFAmod2_promax <- fa(ARQ_cor, nfactors = 2, fm = "pa", rotate = "promax")
AR_EFAmod2_promax
# Oblimin
AR_EFAmod2_oblimin <- fa(ARQ_cor, nfactors = 2, n.obs =nrow(animalrights_na), fm = "pa", rotate = "oblimin")
AR_EFAmod2_oblimin

AR_EFAmod2_promax$communality
```

Diagram function:
```{r}
fa.diagram(AR_EFAmod2_varimax )

fviz_loadnings_with_cor(AR_EFAmod2_oblimin, axes = 1, loadings_above = 0.4)
```

New model building. i extracted ar16, because it didn't have any high correlation in any of the factors, further, by looking at the questions, it might be a causal relationship or be a result of confounding, as it was a religious question, which didn't fit into the themes of the other questions:
```{r}
AR_EFAmod3 <- fa(ARQ_cor[, - c(16)], nfactors = 2, fm = "pa")
AR_EFAmod3_varimax <- fa(ARQ_cor[, - c(16)], nfactors = 2, fm = "pa", rotate = "varimax")
AR_EFAmod3_varimax 

AR_EFAmod3_promax <- fa(ARQ_cor[, - c(16)], nfactors = 2, fm = "pa", rotate = "promax")
AR_EFAmod3_promax

mean(AR_EFAmod3$communality)

print(AR_EFAmod3_promax$loadings, cutoff = 0.3)
fa.diagram(AR_EFAmod3_varimax )
```

I would like to remove ar12 as it seems to be in between the two factors, and the question is about rights, which doesn't nescessarily fit with the other questions: 
```{r}
AR_EFAmod4 <- fa(ARQ_cor[, - c(12,16)], nfactors = 2, fm = "pa")

AR_EFAmod4_promax <- fa(ARQ_cor[, - c(12,16)], nfactors = 2, fm = "pa", rotate = "promax")
AR_EFAmod4_promax

mean(AR_EFAmod4_promax$communality)

print(AR_EFAmod4_promax$loadings, cutoff = 0.3)

AR_EFAmod4_c <- as.data.frame(sort(AR_EFAmod4_promax$communality, decreasing = TRUE))
AR_EFAmod4_c
```

As there seem to be cross loading with ar15, and this question also regards rights and therefore does not fit with other questions, i choose to remove it. It has a very high communality but that might be a result of a causal relationship:
```{r}
AR_EFAmod5 <- fa(ARQ_cor[, - c(12,15,16)], nfactors = 2, fm = "pa")

AR_EFAmod5_promax <- fa(ARQ_cor[, - c(12,15,16)], nfactors = 2, fm = "pa", rotate = "promax")
AR_EFAmod5_promax

mean(AR_EFAmod5_promax$communality)
print(AR_EFAmod5_promax$loadings, cutoff = 0.3)

AR_EFAmod5_c <- as.data.frame(sort(AR_EFAmod5_promax$communality, decreasing = TRUE))
AR_EFAmod5_c

AR_EFAmod5_varimax <- fa(ARQ_cor[, - c(12,15,16)], nfactors = 2, fm = "pa", rotate = "varimax")
AR_EFAmod5_varimax
print(AR_EFAmod5_varimax$loadings, cutoff = 0.3)
# doesn't seem as if varimax works for my analysis. I will continue using promax. I allow some overlapping as some of the questions are close to each other, and all concern animal rights.

fa.diagram(AR_EFAmod5_promax )
fviz_loadnings_with_cor(AR_EFAmod5_promax, axes = 1, loadings_above = 0.4)
fviz_loadnings_with_cor(AR_EFAmod5_promax, axes = 2, loadings_above = 0.4)
```

As i can see in my evolvement of the factors many of the questions are close to each other and overlap. In order to find the underlying factors that shapes the attitudes i would like to go back a few steps and spread out the questions on three factors. Theoretically, this also makes sense:
```{r}
AR_EFAmod3 <- fa(ARQ_cor[, - c(16)], nfactors = 3, fm = "pa")
AR_EFAmod3_varimax <- fa(ARQ_cor[, - c(16)], nfactors = 3, fm = "pa", rotate = "varimax")
AR_EFAmod3_varimax 

AR_EFAmod3_promax <- fa(ARQ_cor[, - c(16)], nfactors = 3, fm = "pa", rotate = "promax")
AR_EFAmod3_promax

mean(AR_EFAmod3$communality)
print(AR_EFAmod3_promax$loadings, cutoff = 0.3)

fa.diagram(AR_EFAmod3_promax)
?fa.diagram
AR_EFAmod3_promax$communality
```
I would like to remove ar1 as it seems to be in between the three factors and is about wild animals and wilderness, which is not a common theme in the questions:
```{r}
AR_EFAmod4 <- fa(ARQ_cor[, - c(1,16)], nfactors = 3, fm = "pa")

AR_EFAmod4_promax <- fa(ARQ_cor[, - c(1,16)], nfactors = 3, fm = "pa", rotate = "promax")
AR_EFAmod4_promax

mean(AR_EFAmod4_promax$communality)

print(AR_EFAmod4_promax$loadings, cutoff = 0.3)

AR_EFAmod4_c <- as.data.frame(sort(AR_EFAmod4_promax$communality, decreasing = TRUE))
AR_EFAmod4_c

fa.diagram(AR_EFAmod4_promax)
```
As there seem to be cross loading with ar15, it has a very high communality but that might be a result of a causal relationship:
```{r}
AR_EFAmod5 <- fa(ARQ_cor[, - c(1,15,16)], nfactors = 3, fm = "pa")
AR_EFAmod5_promax <- fa(ARQ_cor[, - c(1,15,16)], nfactors = 3, fm = "pa", rotate = "promax")
AR_EFAmod5_promax

mean(AR_EFAmod5_promax$communality)

print(AR_EFAmod5_promax$loadings, cutoff = 0.3)

AR_EFAmod5_c <- as.data.frame(sort(AR_EFAmod5_promax$communality, decreasing = TRUE))
AR_EFAmod5_c

AR_EFAmod5_varimax <- fa(ARQ_cor[, - c(12,15,16)], nfactors = 2, fm = "pa", rotate = "varimax")
AR_EFAmod5_varimax
print(AR_EFAmod5_varimax$loadings, cutoff = 0.3)
# doesn't seem as if varimax works for my analysis. I will continue using promax. I allow some overlapping as some of the questions are close to each other, and all concern animal rights.
fa.diagram(AR_EFAmod5_promax )

fviz_loadnings_with_cor(AR_EFAmod5_promax, axes = 1, loadings_above = 0.4)
fviz_loadnings_with_cor(AR_EFAmod5_promax, axes = 2, loadings_above = 0.4)
```
Question ar12 also seems to be creating cross loadings, it is about rights, and it might create some confusion. Even though it has as good loading on both:
```{r}
AR_EFAmod6 <- fa(ARQ_cor[, - c(1,12,15,16)], nfactors = 3, fm = "pa")
AR_EFAmod6

AR_EFAmod6_promax <- fa(ARQ_cor[, - c(1,12,15,16)], nfactors = 3, fm = "pa", rotate = "promax")
AR_EFAmod6_promax

AR_EFAmod6_oblimin <- fa(ARQ_cor[, - c(1,12,15,16)], nfactors = 3, fm = "pa", rotate = "oblimin")
AR_EFAmod6_oblimin

mean(AR_EFAmod6_promax$communality)

print(AR_EFAmod6_promax$loadings, cutoff = 0.3)
print(AR_EFAmod6_oblimin$loadings, cutoff = 0.3)

AR_EFAmod6_c <- as.data.frame(sort(AR_EFAmod6_promax$communality, decreasing = TRUE))
AR_EFAmod6_c

fviz_loadnings_with_cor(AR_EFAmod6_promax, axes = 2, loadings_above = 0.4)
```
I want to remove ar11 now because it is creating a cross loading and not very strong in any of the cases. Theoretically, it is also about insects/pests which is a more specific question and therefore not so easily fit on other questions:
```{r}
AR_EFAmod7 <- fa(ARQ_cor[, - c(1,11,12,15,16)], nfactors = 3, fm = "pa")
AR_EFAmod7

AR_EFAmod7_promax <- fa(ARQ_cor[, - c(1,11,12,15,16)], nfactors = 3, fm = "pa", rotate = "promax")
AR_EFAmod7_promax

mean(AR_EFAmod7_promax$communality)

print(AR_EFAmod7_promax$loadings, cutoff = 0.3)
fa.diagram(AR_EFAmod7_promax)
```
This made the model worse, i want to go back and re-add some of the questions: I am adding ar12 again because theoretically it might be a stronger fit for a factor:
```{r}
AR_EFAmod8 <- fa(ARQ_cor[, - c(1,11,15,16)], nfactors = 3, fm = "pa")
AR_EFAmod8

AR_EFAmod8_promax <- fa(ARQ_cor[, - c(1,11,15,16)], nfactors = 3, fm = "pa", rotate = "promax")
AR_EFAmod8_promax

mean(AR_EFAmod8_promax$communality)

print(AR_EFAmod8_promax$loadings, cutoff = 0.3)
fa.diagram(AR_EFAmod7_promax)
```
It it still struggling, i will therefore try to remove ar15 as well, since it is close to ar12:
```{r}
AR_EFAmod9 <- fa(ARQ_cor[, - c(1,11,16)], nfactors = 3, fm = "pa")
AR_EFAmod9

AR_EFAmod9_promax <- fa(ARQ_cor[, - c(1,11,16)], nfactors = 3, fm = "pa", rotate = "promax")
AR_EFAmod9_promax

mean(AR_EFAmod9_promax$communality)

print(AR_EFAmod9_promax$loadings, cutoff = 0.3)
fa.diagram(AR_EFAmod9_promax)

fviz_loadnings_with_cor(AR_EFAmod9_promax, axes = 1, loadings_above = 0.4)
fviz_loadnings_with_cor(AR_EFAmod9_promax, axes = 2, loadings_above = 0.4)
```
As i am aware that i almost went back to scratch, i now need to figure out which questions will make my new model better.
For my next model, i will be removing ar14 as it is also a radical strong question theoretically and is not strong in any factor:
```{r}
AR_EFAmod10 <- fa(ARQ_cor[, - c(1,11,14,16)], nfactors = 3, fm = "pa")
AR_EFAmod10

AR_EFAmod10_promax <- fa(ARQ_cor[, - c(1,11,14,16)], nfactors = 3, fm = "pa", rotate = "promax")
AR_EFAmod10_promax

mean(AR_EFAmod10_promax$communality)

print(AR_EFAmod10_promax$loadings, cutoff = 0.3)
fa.diagram(AR_EFAmod10_promax)

fviz_loadnings_with_cor(AR_EFAmod10_promax, axes = 1, loadings_above = 0.4) 
fviz_loadnings_with_cor(AR_EFAmod10_promax, axes = 2, loadings_above = 0.4)
```
ar23 fits well with some of the questions in V1, but it is also affecting V2, which is making the model harder to interpret. Furhter, multiple questions are still cross correlated, and removing ar23 might change that:
```{r}
AR_EFAmod11 <- fa(ARQ_cor[, - c(1,11,14,16,23)], nfactors = 3, fm = "pa")
AR_EFAmod11

AR_EFAmod11_promax <- fa(ARQ_cor[, - c(1,11,14,16,23)], nfactors = 3, fm = "pa", rotate = "promax")
AR_EFAmod11_promax

mean(AR_EFAmod11_promax$communality)

print(AR_EFAmod11_promax$loadings, cutoff = 0.3)
fa.diagram(AR_EFAmod11_promax)

fviz_loadnings_with_cor(AR_EFAmod11_promax, axes = 1, loadings_above = 0.4) 
fviz_loadnings_with_cor(AR_EFAmod11_promax, axes = 2, loadings_above = 0.4)
fviz_loadnings_with_cor(AR_EFAmod11_promax, axes = 3, loadings_above = 0.4)

AR_EFAmod11_promax$communality
```

When visualising the factors in fviz i can see that ar8 is an issue when looking at V2 and it is one of the lower communalities in the model. otherwise the other quesions fit together. after rerunning , ar23 was added again as it did not have the wanted effect:
```{r}
AR_EFAmod12 <- fa(ARQ_cor[, - c(1,8,11,14,16)], nfactors = 3, fm = "pa")
AR_EFAmod12

AR_EFAmod12_promax <- fa(ARQ_cor[, - c(1,8,11,14,16)], nfactors = 3, fm = "pa", rotate = "promax")
AR_EFAmod12_promax

mean(AR_EFAmod12_promax$communality)

print(AR_EFAmod12_promax$loadings, cutoff = 0.3)
fa.diagram(AR_EFAmod12_promax)

fviz_loadnings_with_cor(AR_EFAmod12_promax, axes = 1, loadings_above = 0.4) 
fviz_loadnings_with_cor(AR_EFAmod12_promax, axes = 2, loadings_above = 0.4)
fviz_loadnings_with_cor(AR_EFAmod12_promax, axes = 3, loadings_above = 0.4)

AR_EFAmod12_promax$communality
```
Since i am having issues with ar18 being very high in factor 3 and i don't see it fiting in the right factor, and the communality is not super high, i chose to remove it so the variables can spread out more clearly:
```{r}
AR_EFAmod13 <- fa(ARQ_cor[, - c(1,8,11,14,16,18)], nfactors = 3, fm = "pa")
AR_EFAmod13

AR_EFAmod13_promax <- fa(ARQ_cor[, - c(1,8,11,14,16,18)], nfactors = 3, fm = "pa", rotate = "promax")
AR_EFAmod13_promax

mean(AR_EFAmod13_promax$communality)

print(AR_EFAmod13_promax$loadings, cutoff = 0.3)
fa.diagram(AR_EFAmod13_promax)

fviz_loadnings_with_cor(AR_EFAmod13_promax, axes = 1, loadings_above = 0.4) 
fviz_loadnings_with_cor(AR_EFAmod13_promax, axes = 2, loadings_above = 0.4)
fviz_loadnings_with_cor(AR_EFAmod13_promax, axes = 3, loadings_above = 0.4)
```
They are more clearly spread out at the moment, in a way that makes more sense. But there is still a few misfits of questions when using fviz. therefore i will remove one of those strongly correlated. In this case i will remove ar20, as it does not fit into the model:
```{r}
AR_EFAmod14 <- fa(ARQ_cor[, - c(1,8,11,14,16,18,20)], nfactors = 3, fm = "pa")
AR_EFAmod14

AR_EFAmod14_promax <- fa(ARQ_cor[, - c(1,8,11,14,16,18,20)], nfactors = 3, fm = "pa", rotate = "promax")
AR_EFAmod14_promax

mean(AR_EFAmod14_promax$communality)

print(AR_EFAmod14_promax$loadings, cutoff = 0.3)
fa.diagram(AR_EFAmod14_promax)

fviz_loadnings_with_cor(AR_EFAmod14_promax, axes = 1, loadings_above = 0.4) 
fviz_loadnings_with_cor(AR_EFAmod14_promax, axes = 2, loadings_above = 0.4)
fviz_loadnings_with_cor(AR_EFAmod14_promax, axes = 3, loadings_above = 0.4)
```
having removed 7 questions, i am very close. But my V1, which has questions about animal research, also have two questions (ar12 and ar15) that is about animal rights. They make sense when looking at people being against animal research, but i would argue, that they also fit into V2, which has questions about saving animal lives. Therefore, i would like to try and remove one more. In this case, i will be removing ar12, since ar15 also shows up for factor 2 (PA2) and i think it would fit better there:
```{r}
AR_EFAmod15 <- fa(ARQ_cor[, - c(1,8,11,12,14,16,18,20)], nfactors = 3, fm = "pa")
AR_EFAmod15

AR_EFAmod15_promax <- fa(ARQ_cor[, - c(1,8,11,12,14,16,18,20)], nfactors = 3, fm = "pa", rotate = "promax")
AR_EFAmod15_promax

mean(AR_EFAmod15_promax$communality)

print(AR_EFAmod15_promax$loadings, cutoff = 0.3)
fa.diagram(AR_EFAmod15_promax)

fviz_loadnings_with_cor(AR_EFAmod15_promax, axes = 1, loadings_above = 0.4) 
fviz_loadnings_with_cor(AR_EFAmod15_promax, axes = 2, loadings_above = 0.4)
fviz_loadnings_with_cor(AR_EFAmod15_promax, axes = 3, loadings_above = 0.4)

AR_EFAmod15_promax$communality
```
After seeing it not having an effect, i would like to go one step back, and instead focus on another factor. As i can see the rights questions both being relevant for V2 and V1, i would like to keep it out of V3, which is more about animal products. Therefore, i would like to add ar12 back in, but remove ar22:
```{r}
AR_EFAmod16 <- fa(ARQ_cor[, - c(1,8,11,14,16,18,20,22)], nfactors = 3, fm = "pa")
AR_EFAmod16

AR_EFAmod16_promax <- fa(ARQ_cor[, - c(1,8,11,14,16,18,20,22)], nfactors = 3, fm = "pa", rotate = "promax")
AR_EFAmod16_promax

mean(AR_EFAmod16_promax$communality)

print(AR_EFAmod16_promax$loadings, cutoff = 0.3)
fa.diagram(AR_EFAmod16_promax)

fviz_loadnings_with_cor(AR_EFAmod16_promax, axes = 1, loadings_above = 0.4) 
fviz_loadnings_with_cor(AR_EFAmod16_promax, axes = 2, loadings_above = 0.4)
fviz_loadnings_with_cor(AR_EFAmod16_promax, axes = 3, loadings_above = 0.4)
```
After running model 16, i can now see, that ar25, which is about research, has moved into V2, which does not fit my theory. i will go back and add ar22 again and remove ar25 as this is a better fit for my theory:
```{r}
AR_EFAmod17 <- fa(ARQ_cor[, - c(1,8,11,14,16,18,20,25)], nfactors = 3, fm = "pa")
AR_EFAmod17

AR_EFAmod17_promax <- fa(ARQ_cor[, - c(1,8,11,14,16,18,20,25)], nfactors = 3, fm = "pa", rotate = "promax")
AR_EFAmod17_promax

mean(AR_EFAmod17_promax$communality)

AR_EFAmod17_promax$communality

print(AR_EFAmod17_promax$loadings, cutoff = 0.4, sort = TRUE)
fa.diagram(AR_EFAmod17_promax)

Factor1 <- fviz_loadnings_with_cor(AR_EFAmod17_promax, axes = 1, loadings_above = 0.4) + labs(title= "Aganst Using Animals in Research")
Factor2 <- fviz_loadnings_with_cor(AR_EFAmod17_promax, axes = 2, loadings_above = 0.4) + labs(title= "Aganst Animal killing")
Factor3 <- fviz_loadnings_with_cor(AR_EFAmod17_promax, axes = 3, loadings_above = 0.4) + labs(title= "Aganst Animal as Products")

list_factors <- list(Factor1, Factor2, Factor3)
list_factors
```

add model to dataframe:
```{r}
Factors <- factor.scores(animalrights_na[,-c(1,8,11,14,16,18,20,25,29,30,31)],AR_EFAmod17_promax)$scores
colnames(Factors) <-  c("research", "animal_killing","product_usage")

view(Factors)

final_data <- cbind(animalrights_na[,-c(1,8,11,14,16,18,20,25)],Factors)

view(final_data)
```

now i want to make a linear model predicting how conservative or liberal the participants are:
```{r}
plot(final_data$liberal,final_data$research)

plot(final_data$liberal, final_data$animal_killing)

predict_liberal <-  lm(liberal ~ research + animal_killing + product_usage, data= final_data)

summary(predict_liberal)

vif(predict_liberal) # ikke multicollinearitet
```

