
Lab assignment 4
Majbritt Lindholt Stenbøg

packages:
```{r}
library(psychTools)
library(lavaan) 
library(semPlot)
library(tidyverse) 
library(CompQuadForm)
library(ICS)
```

Dataset:
```{r}
my_data = holzinger.swineford

view(my_data)
```

Task 1:

Model A:
```{r}
model_A <- 
"
# Measurement model
Vis_per =~ t01_visperc + t02_cubes + t03_frmbord + t04_lozenges
Verbal_a =~ t06_paracomp + t07_sentcomp + t09_wordmean
Pros_spd =~ t10_addition + t12_countdot + t13_sccaps 

"
```

Fit the model:
```{r}
fitmodelA <- sem(model_A, data=my_data)

semPaths(fitmodelA)
```

Estimate normal distribution:
```{r}
summary(fitmodelA)

mvnorm.kur.test(my_data[,c("t01_visperc", "t02_cubes", "t03_frmbord", "t04_lozenges", "t06_paracomp",  "t07_sentcomp", "t09_wordmean","t10_addition", "t12_countdot", "t13_sccaps")]) 
# significant

mvnorm.skew.test(my_data[,c("t01_visperc", "t02_cubes", "t03_frmbord", "t04_lozenges", "t06_paracomp",  "t07_sentcomp", "t09_wordmean","t10_addition", "t12_countdot", "t13_sccaps")]) 
# significant
```

So it violates normal distribution. Now i will fix this by adding robust estimates:
```{r}
fitmodelA_MLM <- sem(model_A, data=my_data, estimator = "MLM")

summary(fitmodelA_MLM, fit.measures= T, rsquare = T, standardized = T)

semPaths(fitmodelA_MLM, whatLabels = "est", label.scale =0, edge.label.cex =0.8)
```

Task 2:
model B:
```{r}
model_B <- 
"
# Measurement model
Vis_per =~ t01_visperc + t02_cubes + t03_frmbord + t04_lozenges
Verbal_a =~ t06_paracomp + t07_sentcomp + t09_wordmean
Pros_spd =~ t10_addition + t12_countdot + t13_sccaps 
# correlation
t10_addition ~~ t12_countdot
"
```

now model B was created. I will run the same code for model B in order to compare it with model A:
```{r}
fitmodelB <- sem(model_B, data=my_data)

semPaths(fitmodelB)
```

```{r}
fitmodelB_MLM <- sem(model_B, data=my_data, estimator = "MLM")

summary(fitmodelB_MLM, fit.measures = T, rsquare = T)
```


which manifest variable is least influenced:
```{r}
summary(fitmodelB_MLM, fit.measures = T, ci =95, rsquare = T,standardized = T)
```

If we want to visualise this with a path diagram:
```{r}
semPaths(fitmodelB_MLM,residuals = T, whatLabels = "std",label.scale =0, edge.label.cex =0.8)
```

task 3:
```{r}
indirect_eff <- 0.23*0.38
direct_eff <-  0.31
indirect_eff + direct_eff
indirect_eff
```




